Blame and Coercion: Together Again for the Second Time

### 2019-10-04 paper.tex
blame calculus

\begin{figure}
	\[
	\begin{array}{lclr}
	\stxrule{S,T}{types}{
		\star \mid{}
		P
	}
	\stxrule{P,Q}{pre-types}{
		\iota \mid{}
		T_1 \rightarrow T_2 \mid{}
		T_1 \times T_2 \mid{}
		T_1 \plus T_2 \mid{}
	}
%	\stxrule{x,y,z}{variables}{\dots}
%	\stxrule{l}{labels}{\dots}
%	\stxrule{k}{constants}{sole}
	\stxrule{e}{terms}{
		x \mid{}
		\mathtt{tt} \mid{}
		\lambda^{S\rightarrow{}T}x.t \mid{}
		e_1 \; e_2 \mid{}
		\langle T \Leftarrow^l S \rangle t \mid{}
		\mathtt{blame} \; l
	}
	\stxrulecont{
		\mathtt{cons}\; e_1 \; e_2 \mid{}
		\mathtt{car}\; e \mid{}
		\mathtt{cdr}\; t
	}
	\stxrulecont{
		\mathtt{inl} \; e \mid{}
		\mathtt{inr} \; e \mid{}
		\mathtt{case}\; e_1 \; e_2 \; e_3
	}
	\stxrule{o}{observations}{
		\mathtt{tt} \mid{}
		\mathtt{fun}\mid{}
		\mathtt{cons}\mid{}
		\mathtt{inl}\mid{}
		\mathtt{inr}\mid{}
		\mathtt{blame}\; l
	}
	\end{array}
	\]

@@

\figref{fig:blame-static} defines the form of blame calculus and its static 
semantics. It is little changed from previous definitions. 
The dynamic semantics of Blame Calculi depend on blame strategies, so we defer 
them to sub-sections.

Blame Calculus is based on Simply Typed Lambda Calculus with sum types and 
product types ($ \mathtt{STLC+} $). 
Let $ S,T $ range over types. A type is either the dynamic type $ \star $
(a.k.a. $ \mathtt{Dyn} $, $ \mathbb{?} $, or $ \mathtt{Unknown} $), 
or a pre-type. 
Let $ P,Q $ range over pre-types. Every pre-type is a type with a 
traditional type constructor at the top.

@@

One might be tempted to conclude that inconsistency is the root of all runtime 
type errors. In the setting of gradual typing, however, runtime type-checking 
is usually gradual as well. 
$ S \smile T $ reads ``$ S $ and $ T $ are shallowly-consistent''. Two types 
are shallowly consistent if one of them is $ \star $, or they have the same 
top-most type constructor. Shallow-inconsistency is the root of all runtime 
type error -- Casting a value to a shallowly inconsistent type would 
lead to an runtime type error immediately. Shallow-consistency is also 
reflexive, symmetric, and not transitive.

Let $ e $ ranges over terms. Unlike $ \mathtt{STLC+} $, we annotate the 
co-domain of lambda abstractions explicitly. Besides, we add casts and blames.

Let $ o $ ranges over observations.

@@

We describe the dynamic semantics of Blame Calculus with the CEK machine 
(\citet{felleisen1986control}).

>>>>

\begin{figure}
	\[
	\begin{array}{lclr}
	\stxrule{S,T}{types}{
		\star \mid{}
		P
	}
	\stxrule{P,Q}{pre-types}{
		\iota \mid{}
		T_1 \rightarrow T_2 \mid{}
		T_1 \times T_2 \mid{}
		T_1 \plus T_2
	}
%	\stxrule{x,y,z}{variables}{\dots}
%	\stxrule{l}{labels}{\dots}
%	\stxrule{k}{constants}{sole}
	\stxrule{e}{terms}{
		x \mid{}
		\mathtt{tt} \mid{}
		\lambda^{S\rightarrow{}T}x.t \mid{}
		e_1 \; e_2
	}
	\stxrulecont{
		\mathtt{cons}\; e_1 \; e_2 \mid{}
		\mathtt{car}\; e \mid{}
		\mathtt{cdr}\; t
	}
	\stxrulecont{
		\mathtt{inl} \; e \mid{}
		\mathtt{inr} \; e \mid{}
		\mathtt{case}\; e_1 \; e_2 \; e_3
	}
	\stxrulecont{
		\langle T \Leftarrow^l S \rangle t \mid{}
		\mathtt{blame} \; l
	}
	\stxrule{o}{observations}{
		\mathtt{tt} \mid{}
		\mathtt{fun}\mid{}
		\mathtt{cons}\mid{}
		\mathtt{inl}\mid{}
		\mathtt{inr}\mid{}
		\mathtt{blame}\; l
	}
	\end{array}
	\]

@@

\figref{fig:blame-static} defines the syntax of blame calculus and its static 
semantics. It is little changed from previous definitions. 
The dynamic semantics of Blame Calculi depend on blame strategies, so we defer 
them to sub-sections.

Blame Calculus is based on Simply Typed Lambda Calculus with sum types and 
product types ($ \mathtt{STLC+} $). 
Let $ S,T $ range over types. A type is either the dynamic type $ \star $
(a.k.a. $ \mathtt{Dyn} $, $ \mathbb{?} $, or $ \mathtt{Unknown} $), 
or a pre-type. 
Let $ P,Q $ range over pre-types. Every pre-type is a type with a type 
constructor at the top. For simplicity, we only have one base type $ \iota $, 
the unit type. Other pre-types are functions, products, and sums.

@@

One might be tempted to conclude that inconsistency is the root of all runtime 
type errors (blames). In the setting of gradual typing, however, runtime 
type-checking is usually gradual as well. 
$ S \smile T $ reads ``$ S $ and $ T $ are shallowly-consistent''. Two types 
are shallowly consistent if one of them is $ \star $, or they have the same 
top-most type constructor. Shallow-inconsistency is the root of all blames -- 
casting a value to a shallowly inconsistent type leads to a blame immediately. 
Shallow-consistency is also reflexive, symmetric, and not transitive.

Let $ e $ ranges over terms, including all terms from $ \mathtt{STLC+} $, 
casts, and blames. Unlike $ \mathtt{STLC+} $, we annotate the 
co-domain of lambda abstractions explicitly. 
The end of the figure shows the extra typing rules.

Let $ o $ ranges over observations. They are what would be observed if a 
program terminates. Observations include all constructors and blames.

@@

We describe the dynamic semantics of \lazyD{} Blame Calculus with the CEK 
machine (\citet{felleisen1986control}).
